---
title: "Common distros"
author: "Yevgeny V.Yorkhov"
date: "04/19/2015"
output: html_document
---

## Normal Distribution

If Z is ***standard normal*** and X is ***normal***:

Z = (X - mu) / sigma ~ N(0,1)

and

X = mu + sigma*Z ~ N(mu, sigma^2)

### Quantiles
* Approximately 68%, 95%, 99% of the normal density lies within 1, 2 and 3 standard deviations from the mean (mu), respectively
* -1.28, -1.645, -1,96 and -2.33 are 10th, 5th, 2.5th and 1st percentiles of the standard normal distribution respectively
* By symmetry 1.28, 1.645, 1.96 and 2.33 are the 90th, 95th, 97.5th and 99th percentiles of the standard normal distribution respectively

### Examples
#### 1. What is 95th quantile of normal distribution
```{r, quantile95}
mu = 0
sd = 1
qnorm(.95, mean = mu, sd=sd)
```

#### 2. Ad clicks 1
Assume that the number of daily ad clicks for a company is (approximately) normally distributed with a mean of 1020 and standard deviation of 50. What is the probability of getting more than 1160 clicks a day?
```{r, adclicks1}
pnorm(1160, mean=1020, sd=50, lower.tail=FALSE)
# or we can see that 1160 is 2.8 standard deviation from the mean
# hence
pnorm(2.8, lower.tail=FALSE)
```
#### 2. Ad clicks 2
Assume that the number of daily ad clicks for a company is (approximately) normally distributed with a mean of 1020 and standard deviation of 50. What number of daily ad clicks would represent the one 75% of days have fewer clicks (assuming days are independent and identically distributed)?
```{r,adclocks2}
qnorm(.75, mean=1020, sd=50)
```

### Plots
```{r}
plot(rnorm(1000, mean=0, sd=1))
hist(rnorm(1000, mean=0, sd=1))
```

## Binomial distribution
### Bernoulli distribution
`P(X=x) = p^x*(1-p)^(1-x)`

* mean = p
* variance = p(1-p)

### Binomial trials
In specific, let X1, X2, ... Xn be iid Bernoulli (p)
then 

`X = sum(from i, till n) Xi ` 

is a binomial random distribution.

Binomial PMF is:

`P(X=x) = choose(n,x) * p^x * (1-p)^(n-x)`

### Examples
Suppose a friend has 8 children (oh my!), 7 of which are girls and none are twins
If each gender has probability 0.5 for each birth, what is the probability of getting
7 or more girls out of 8 birth?

```{r, binomial1}
choose(8, 7) * 0.5^8 + choose(8, 8) * 0.5^8
# or
pbinom(6, size = 8, prob = 0.5, lower.tail = FALSE)
```

### Plots
```{r, plots_binomial}
plot(rbinom(1000, size=10, prob=0.70))
hist(rbinom(1000, size=10, prob=0.70))
```

