---
title: "T_confidence_interval"
author: "Yevgeny V.Yorkhov"
date: "04/25/2015"
output: html_document
---

# T-confidence interval

* Invented by William Gosset (under the pseudonym "Student") in 1908
* has thicker tails than Normal
* Is indexed by ***degrees of freedom*** (df); gets more like a standard normal as df gets larger
* it assumes that underlying data are iid Gaussian with the result that (X - mu) / (S/sqrt(n)) follows Gosset's T-distribution with ***n-1*** degrees of freedom
* Interval is ***X+-t(n-1)S/sqrt(n)*** is the relevant quantile

This code illustrates that T-distribution gets more like a standard normal as ***df*** gets larger
```{r, t_vs_normal}
k <- 1000
xvals <- seq(-5, 5, length = k)
myplot <- function(df){
 d <- data.frame(y = c(dnorm(xvals), dt(xvals, df)),
 x = xvals,
 dist = factor(rep(c("Normal", "T"), c(k,k))))
 g <- ggplot(d, aes(x = x, y = y))
 g <- g + geom_line(size = 2, aes(colour = dist))
 g
}
#manipulate(myplot(mu), mu = slider(1, 20, step = 1))

```

You can also embed plots, for example:

```{r, echo=FALSE}
pvals <- seq(.5, .99, by = .01)
myplot2 <- function(df){
  d <- data.frame(n= qnorm(pvals),t=qt(pvals, df),
         p = pvals)
  g <- ggplot(d, aes(x= n, y = t))

  g <- g + geom_abline(size = 2, col = "lightblue")
  g <- g + geom_line(size = 2, col = "black")
  g <- g + geom_vline(xintercept = qnorm(0.975))
  g <- g + geom_hline(yintercept = qt(0.975, df))
  g
}
manipulate(myplot2(df), df = slider(1, 20, step = 1))
```

In R data(sleep) brings up the sleep data originally analyzed in Gosset's Biometrika paper,
which shows the increase in hours for 10 patients on two soporific drugs. R treats the data as two
groups rather than paired.
```{r}
data(sleep)
head(sleep)

g1 <- sleep$extra[1 : 10]; g2 <- sleep$extra[11 : 20]
difference <- g2 - g1
mn <- mean(difference); s <- sd(difference); n <- 10

mn + c(-1, 1) * qt(.975, n-1) * s / sqrt(n)
t.test(difference)
t.test(g2, g1, paired = TRUE)
t.test(extra ~ I(relevel(group, 2)), paired = TRUE, data = sleep)

```


```{r}
library(datasets); data(ChickWeight); library(reshape2)
##define weight gain or loss
wideCW <- dcast(ChickWeight, Diet + Chick ~ Time, value.var = "weight")
names(wideCW)[-(1 : 2)] <- paste("time", names(wideCW)[-(1 : 2)], sep = "")
library(dplyr)

wideCW <- mutate(wideCW,
  gain = time21 - time0
  )

```

